### 2012-10-01 manuscript/cloudbreak.tex
manuscript

\documentclass[12pt]{article}
\usepackage{graphics}
\usepackage{subfigure}

\begin{document}

\title{A MapReduce Algorithm for Detecting Genomic Structural Variation from Complete Paired-End Sequence Data Sets}

\author{
Christopher Whelan\ Kemal S\"onmez\ \\
Biomedical Engineering \\
Oregon Health and Science University \\
20000 NW Walker Road, Beaverton, OR 97006, USA
}

\date{\today}

\maketitle

\begin{abstract}
This is the paper's abstract \ldots
\end{abstract}

\section{Introduction}
Genomic structural variants are important..

Difficulty of detection.. Lack of concordance of current algorithms...

Typical algorithms use only unambiguous discordant read pair mappings...

Hadoop/MapReduce as framework for processing complete data sets

\section{Methods}\label{Methods}

MapReduce algorithm description

Local deletion scores

\section{Results}\label{results}

Simulated data

1000 Genomes Data

AML Data set

\section{Discussion}\label{discussion}

\section{Conclusions}\label{conclusions}

\bibliographystyle{abbrv}
\bibliography{cloudbreak}

\end{document}