### 2013-02-27 manuscript/NA18507preds.table.brew.tex
brew latex tables

% latex table generated in R 2.15.2 by xtable 1.7-0 package
% Tue Jan  1 22:36:02 2013
\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
 & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
Total Number & & & <%= format(totals['40-100bp'], big.mark=",") %> & <%= format(totals['101-250bp'], big.mark=",") %> & <%= format(totals['251-500bp'], big.mark=",") %> & <%= format(totals['501-1000bp'], big.mark=",") %> & <%= format(totals['> 1000bp'], big.mark=",") %> \\
  \hline
Cloudbreak & <%= printPredTableValueFromList('Cloudbreak', precision) %> & <%= printPredTableValueFromList('Cloudbreak', recall) %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '> 1000bp') %> \\ 
Breakdancer & <%= printPredTableValueFromList('Breakdancer', precision) %> & <%= printPredTableValueFromList('Breakdancer', recall) %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Breakdancer', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '> 1000bp') %> \\  
  GASVPro & <%= printPredTableValueFromList('GASVPro', precision) %> & <%= printPredTableValueFromList('GASVPro',recall) %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'GASVPro', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'GASVPro', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '> 1000bp') %> \\ 
  DELLY & <%= printPredTableValueFromList('DELLY', precision) %> & <%= printPredTableValueFromList('DELLY', recall) %> & <%= printPredTableCell(preds, exPreds, 'DELLY', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'DELLY', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'DELLY', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY', '> 1000bp') %> \\ 
  Pindel & <%= printPredTableValueFromList('Pindel', precision) %> & <%= printPredTableValueFromList('Pindel', recall) %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Pindel', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Pindel', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '> 1000bp') %> \\ 
   \hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of each tool on the NA18507 data, as well as the number of deletions in each size class found. The same cutoffs were used as for the simulated data as in Table \ref{chr2preds}, adjusted for the difference in coverage from 30X to 37X.}
\label{NA18507preds}
\end{table}