### 2013-02-28 manuscript/chr2InsertionPredsBySize.table.brew.tex
everything is now up-to-speed with auto R output

\begin{table}[t]
\begin{center}
\begin{tabular}{rrrrrr}
  \hline
 & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
 Total Number & <%= totals['40-100bp'] %> &  <%= totals['101-250bp'] %> & <%= totals['251-500bp'] %> &  <%= totals['501-1000bp'] %> & <%= totals['> 1000bp'] %>\\ 
  \hline
  Cloudbreak  & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '> 1000bp') %> \\ 
  Breakdancer & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Breakdancer', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '> 1000bp') %> \\ 
  Pindel      & <%= printPredTableCell(preds, exPreds, 'Pindel', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Pindel', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Pindel', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '> 1000bp') %>\\ 
   \hline
\end{tabular}
\end{center}
\caption{The number of simulated insertions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivity, as well as the number of those insertions that were discovered exclusively by each tool (in parentheses). The total number of insertions in each size class in the true set of insertions is shown in the second row of the header.}
\label{chr2InsertionPreds}
\end{table}