### 2013-02-28 manuscript/deletionGenotypeAccuracy.table.brew.tex

\begin{table}
\begin{center}
\begin{tabular}{r|r|rr|rr|}
\multicolumn{2}{c}{}  & \multicolumn{4}{c}{Actual Genotypes} \\
\multicolumn{2}{c}{}  & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\cline{3-6}
\multicolumn{2}{c|}{} &  Homozygous & Heterozygous & Homozygous & Heterozygous \\ 
\cline{2-6}
\multirow{2}{*}{\shortstack{Predicted \\ Genotypes}} & Homozygous & <%= chr2HapCM["2", "TRUE"] %> & <%= chr2HapCM["1", "TRUE"] %> &  <%= NA18507HapCM["2", "TRUE"] %> & <%= NA18507HapCM["1", "TRUE"] %> \\
 & Heterozygous & <%= chr2HapCM["2", "FALSE"] %> & <%= chr2HapCM["1", "FALSE"] %> &  <%= NA18507HapCM["2", "FALSE"] %> & <%= NA18507HapCM["1", "FALSE"] %> \\
\cline{2-6}
\end{tabular}
\end{center}
\caption{Confusion matrices for the predicted genotype of deletions found by Cloudbreak on both the simulated and NA18507 data sets.}
\label{genotypeaccuracy}
\end{table}