### 2013-02-28 manuscript/insertionRepmask.table.brew.tex
finish adding first cut at insertion accurracy to manuscript

\begin{table}[b]
\begin{center}
\begin{tabular}{rrr|rr}
 & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\hline
 &  Non-repeat & Repeat  &  Non-repeat & Repeat \\ 
 Total Number & <%= chr2Totals['0'] %> & <%= chr2Totals['1'] %> & <%= NA18507Totals['0'] %> & <%= NA18507Totals['1'] %> \\ 
  \hline
  Cloudbreak  & <%= printPredTableCell(chr2Preds, chr2ExPreds, 'Cloudbreak', '0') %> & <%= printPredTableCell(chr2Preds, chr2ExPreds, 'Cloudbreak', '1') %> & <%= printPredTableCell(NA18507Preds, NA18507ExPreds, 'Cloudbreak', '0') %> & <%= printPredTableCell(NA18507Preds, NA18507ExPreds, 'Cloudbreak', '1') %> \\ 
  Breakdancer & <%= printPredTableCell(chr2Preds, chr2ExPreds, 'Breakdancer', '0') %> & <%= printPredTableCell(chr2Preds, chr2ExPreds, 'Breakdancer', '1') %> & <%= printPredTableCell(NA18507Preds, NA18507ExPreds, 'Breakdancer', '0') %> & <%= printPredTableCell(NA18507Preds, NA18507ExPreds, 'Breakdancer', '1') %> \\
  Pindel      & <%= printPredTableCell(chr2Preds, chr2ExPreds, 'Pindel', '0') %> & <%= printPredTableCell(chr2Preds, chr2ExPreds, 'Pindel', '1') %> & <%= printPredTableCell(NA18507Preds, NA18507ExPreds, 'Pindel', '0') %> & <%= printPredTableCell(NA18507Preds, NA18507ExPreds, 'Pindel', '1') %> \\ 
   \hline
\end{tabular}
\end{center}
\caption{Detected insertions in the simulated and NA18507 data sets identified by each tool, broken down by whether the insertion occurs in a RepeatMasker-annotated element.}
\label{insertionRepmaskpreds}
\end{table}