### 2013-04-05 manuscript/chr2DeletionPredsMaxSensitivity.table.brew.tex

\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
                        & Prec. & Recall & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
 Total Number &          &           & <%= totals['40-100bp'] %> &  <%= totals['101-250bp'] %> & <%= totals['251-500bp'] %> &  <%= totals['501-1000bp'] %> & <%= totals['> 1000bp'] %>\\ 
  \hline
  Cloudbreak    &  <%= printPredTableValueFromList('Cloudbreak', precision) %> & <%= printPredTableValueFromList('Cloudbreak', recall) %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '> 1000bp') %> \\ 
  Breakdancer   &  <%= printPredTableValueFromList('Breakdancer', precision) %> & <%= printPredTableValueFromList('Breakdancer', recall) %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Breakdancer', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '> 1000bp') %> \\ 
  GASVPro        & <%= printPredTableValueFromList('GASVPro', precision) %> & <%= printPredTableValueFromList('GASVPro', recall) %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'GASVPro', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'GASVPro', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '> 1000bp') %> \\ 
  DELLY-RP           & <%= printPredTableValueFromList('DELLY', precision) %> & <%= printPredTableValueFromList('DELLY', recall) %> & <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '> 1000bp') %> \\ 
  DELLY-SR           & <%= printPredTableValueFromList('DELLYSR', precision) %> & <%= printPredTableValueFromList('DELLYSR', recall) %> & <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '> 1000bp') %> \\ 
  Pindel           & <%= printPredTableValueFromList('Pindel', precision) %> & <%= printPredTableValueFromList('Pindel', recall) %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Pindel', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Pindel', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '> 1000bp') %>\\ 
  MoDIL           & <%= printPredTableValueFromList('MoDIL', precision) %>  & <%= printPredTableValueFromList('MoDIL', recall) %> & <%= printPredTableCell(preds, exPreds, 'MoDIL', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'MoDIL', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'MoDIL', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'MoDIL', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'MoDIL', '> 1000bp') %>\\ 
   \hline
\end{tabular}
\end{center}
\caption{The number of simulated deletions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivy, as well as the number of those deletions that were discovered exclusively by each tool (in parentheses). The total number of deletions in each size class in the true set of deletions is shown in the second row of the header.}
\label{chr2DeletionPredsMaxSensitivity}
\end{table}