### 2013-05-29 manuscript/NA18507DeletionAndInsertionPredsBySize.table.brew.tex
updates to NM manuscript

\begin{table}
\begin{center}
\begin{tabular}{r|rrr|rrrrr}
  \cline{2-9}
&  & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
\hline
\multirow{6}{*}{\begin{sideways}Deletions\end{sideways}} & Total Number & & & <%= format(totalDels['40-100bp'], big.mark=",") %> & <%= format(totalDels['101-250bp'], big.mark=",") %> & <%= format(totalDels['251-500bp'], big.mark=",") %> & <%= format(totalDels['501-1000bp'], big.mark=",") %> & <%= format(totalDels['> 1000bp'], big.mark=",") %> \\
  \hline
\cline{2-9}
& Cloudbreak & <%= printPredTableValueFromList('Cloudbreak', delPrecision) %> & <%= printPredTableValueFromList('Cloudbreak', delRecall) %> & <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '> 1000bp') %> \\ 
& Breakdancer & <%= printPredTableValueFromList('Breakdancer', delPrecision) %> & <%= printPredTableValueFromList('Breakdancer', delRecall) %> & <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '> 1000bp') %> \\  
&  GASVPro & <%= printPredTableValueFromList('GASVPro', delPrecision) %> & <%= printPredTableValueFromList('GASVPro',delRecall) %> & <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '> 1000bp') %> \\ 
&  DELLY-RP & <%= printPredTableValueFromList('DELLY', delPrecision) %> & <%= printPredTableValueFromList('DELLY', delRecall) %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '> 1000bp') %> \\ 
&  DELLY-SR & <%= printPredTableValueFromList('DELLYSR', delPrecision) %> & <%= printPredTableValueFromList('DELLYSR', delRecall) %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '> 1000bp') %> \\ 
&  Pindel & <%= printPredTableValueFromList('Pindel', delPrecision) %> & <%= printPredTableValueFromList('Pindel', delRecall) %> & <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '> 1000bp') %> \\ 
\hline
\multirow{4}{*}{\begin{sideways}Insertions\end{sideways}} & Total Number & & & <%= format(totalIns['40-100bp'], big.mark=",") %> & <%= format(totalIns['101-250bp'], big.mark=",") %> & <%= format(totalIns['251-500bp'], big.mark=",") %> & <%= format(totalIns['501-1000bp'], big.mark=",") %> & 0 \\
\cline{2-9}
& Cloudbreak & <%= printPredTableValueFromList('Cloudbreak', insPrecision) %> & <%= printPredTableValueFromList('Cloudbreak', insRecall) %> & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '40-100bp') %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '101-250bp') %> &  <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '251-500bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '501-1000bp') %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '> 1000bp') %>  \\ 
& Breakdancer & <%= printPredTableValueFromList('Breakdancer', insPrecision) %> & <%= printPredTableValueFromList('Breakdancer', insRecall) %> & <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '40-100bp') %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '101-250bp') %> &  <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '251-500bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '501-1000bp')  %> & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '> 1000bp') %> \\  
&  Pindel & <%= printPredTableValueFromList('Pindel', insPrecision) %> & <%= printPredTableValueFromList('Pindel', insRecall) %> & <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '40-100bp') %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '101-250bp') %> &  <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '251-500bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '501-1000bp')  %> &  <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '> 1000bp')  %> \\ 
\hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of deletions and insertions for each tool on the NA18507 data, as well as the number of variants found in each size class found. Exclusive predictions are in parentheses. For deletions, the same cutoffs were used as for the simulated data as in Table \ref{chr2DeletionPredsFDR10}, adjusted for the difference in coverage from 30X to 37X. For insertions, the maximum sensitivity cutoff was used.}
\label{NA18507DeletionAndInsertionPreds}
\end{table}