### 2013-06-02 manuscript/NA18507DeletionPredsBySize.table.brew.tex

\begin{table}[t]
\begin{center}
\begin{tabular}{rrr|rrrrr}
  \hline
 & Prec. & Recall & 40-100bp & 101-250bp & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
Total Number & & & <%= format(totals['40-100bp'], big.mark=",") %> & <%= format(totals['101-250bp'], big.mark=",") %> & <%= format(totals['251-500bp'], big.mark=",") %> & <%= format(totals['501-1000bp'], big.mark=",") %> & <%= format(totals['> 1000bp'], big.mark=",") %> \\
  \hline
Cloudbreak & <%= printPredTableValueFromList('Cloudbreak', precision) %> & <%= printPredTableValueFromList('Cloudbreak', recall) %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Cloudbreak', '> 1000bp') %> \\ 
Breakdancer & <%= printPredTableValueFromList('Breakdancer', precision) %> & <%= printPredTableValueFromList('Breakdancer', recall) %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Breakdancer', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Breakdancer', '> 1000bp') %> \\  
  GASVPro & <%= printPredTableValueFromList('GASVPro', precision) %> & <%= printPredTableValueFromList('GASVPro',recall) %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'GASVPro', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'GASVPro', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'GASVPro', '> 1000bp') %> \\ 
  DELLY-RP & <%= printPredTableValueFromList('DELLY', precision) %> & <%= printPredTableValueFromList('DELLY', recall) %> & <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY-RP', '> 1000bp') %> \\ 
  DELLY-SR & <%= printPredTableValueFromList('DELLYSR', precision) %> & <%= printPredTableValueFromList('DELLYSR', recall) %> & <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'DELLY-SR', '> 1000bp') %> \\ 
  Pindel & <%= printPredTableValueFromList('Pindel', precision) %> & <%= printPredTableValueFromList('Pindel', recall) %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '40-100bp') %>  & <%= printPredTableCell(preds, exPreds, 'Pindel', '101-250bp') %> &  <%= printPredTableCell(preds, exPreds, 'Pindel', '251-500bp') %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '501-1000bp') %> & <%= printPredTableCell(preds, exPreds, 'Pindel', '> 1000bp') %> \\ 
   \hline
\end{tabular}
\end{center}
\caption{The precision and recall with respect to the gold standard set of each tool on the NA18507 data, as well as the number of deletions in each size class found. The same cutoffs were used as for the simulated data as in Table \ref{chr2DeletionPredsFDR10}, adjusted for the difference in coverage from 30X to 37X.}
\label{NA18507DeletionPreds}
\end{table}