### 2013-12-13 manuscript/chr2AccuracyByWindowSize.table.brew.tex

\begin{table}
\begin{center}
\begin{tabular}{r|rrrrr}
 \hline
 Window Size & Calls & True Positives & Precision & Recall & F \\ 
 \hline
   <%= formatNumber(windowTable[1,'Resolution']) %> & <%= formatNumber(windowTable[1,'Calls']) %> & <%= formatNumber(windowTable[1,'TP']) %> & <%= formatNumber(windowTable[1,'Precision']) %> & <%= formatNumber(windowTable[1,'Recall']) %> & <%= formatNumber(windowTable[1,'F']) %> \\ 
   <%= formatNumber(windowTable[2,'Resolution']) %> & <%= formatNumber(windowTable[2,'Calls']) %> & <%= formatNumber(windowTable[2,'TP']) %> & <%= formatNumber(windowTable[2,'Precision']) %> & <%= formatNumber(windowTable[2,'Recall']) %> & <%= formatNumber(windowTable[2,'F']) %> \\ 
   <%= formatNumber(windowTable[3,'Resolution']) %> & <%= formatNumber(windowTable[3,'Calls']) %> & <%= formatNumber(windowTable[3,'TP']) %> & <%= formatNumber(windowTable[3,'Precision']) %> & <%= formatNumber(windowTable[3,'Recall']) %> & <%= formatNumber(windowTable[3,'F']) %> \\ 
   <%= formatNumber(windowTable[4,'Resolution']) %> & <%= formatNumber(windowTable[4,'Calls']) %> & <%= formatNumber(windowTable[4,'TP']) %> & <%= formatNumber(windowTable[4,'Precision']) %> & <%= formatNumber(windowTable[4,'Recall']) %> & <%= formatNumber(windowTable[4,'F']) %> \\ 
   <%= formatNumber(windowTable[5,'Resolution']) %> & <%= formatNumber(windowTable[5,'Calls']) %> & <%= formatNumber(windowTable[5,'TP']) %> & <%= formatNumber(windowTable[5,'Precision']) %> & <%= formatNumber(windowTable[5,'Recall']) %> & <%= formatNumber(windowTable[5,'F']) %> \\ 
   <%= formatNumber(windowTable[6,'Resolution']) %> & <%= formatNumber(windowTable[6,'Calls']) %> & <%= formatNumber(windowTable[6,'TP']) %> & <%= formatNumber(windowTable[6,'Precision']) %> & <%= formatNumber(windowTable[6,'Recall']) %> & <%= formatNumber(windowTable[6,'F']) %> \\  
\end{tabular}
\end{center}
\caption{Accuracy measures for Cloudbreak on the Chromosome 2 simulated data set using different choices of window size. For each window size the same Cloudbreak score threshold was used (1.98).}
\label{chr2AccuracyByWindowSize}
\end{table}