### 2013-12-13 manuscript/chr2DeletionAndInsertionPredsMaxSensitivity.table.brew.tex

\begin{table}[t]
\begin{center}
\begin{tabular}{r|rrrr|rrrrr}
  \cline{2-10}
   &                     & Prec. & Recall & F1 & 40-100bp  & 101-250bp  & 251-500bp & 501-1000bp & $>$ 1000bp \\ 
\hline
\multirow{7}{*}{\begin{sideways}Deletions\end{sideways}} & Total Number &          &           & &  <%= totalDels['40-100bp'] %> &  <%= totalDels['101-250bp'] %> & <%= totalDels['251-500bp'] %> &  <%= totalDels['501-1000bp'] %> & <%= totalDels['> 1000bp'] %>\\ 
  \hline
\cline{2-10}
&  Cloudbreak    &  <%= printPredTableValueFromList('Cloudbreak', delPrecision) %> & <%= printPredTableValueFromList('Cloudbreak', delRecall) %> & <%= printPredTableValueFromList('Cloudbreak', delf1) %> & <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Cloudbreak', '> 1000bp') %> \\ 
&  Breakdancer   &  <%= printPredTableValueFromList('Breakdancer', delPrecision) %> & <%= printPredTableValueFromList('Breakdancer', delRecall) %> & <%= printPredTableValueFromList('Breakdancer', delf1) %> & <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Breakdancer', '> 1000bp') %> \\ 
&  GASVPro        & <%= printPredTableValueFromList('GASVPro', delPrecision) %> & <%= printPredTableValueFromList('GASVPro', delRecall) %> & <%= printPredTableValueFromList('GASVPro', delf1) %> & <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'GASVPro', '> 1000bp') %> \\ 
&  DELLY-RP           & <%= printPredTableValueFromList('DELLY', delPrecision) %> & <%= printPredTableValueFromList('DELLY', delRecall) %> & <%= printPredTableValueFromList('DELLY', delRecall) %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-RP', '> 1000bp') %> \\ 
&  DELLY-SR           & <%= printPredTableValueFromList('DELLYSR', delPrecision) %> & <%= printPredTableValueFromList('DELLYSR', delRecall) %> &  <%= printPredTableValueFromList('DELLYSR', delf1) %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'DELLY-SR', '> 1000bp') %> \\ 
&  Pindel           & <%= printPredTableValueFromList('Pindel', delPrecision) %> & <%= printPredTableValueFromList('Pindel', delRecall) %> &  <%= printPredTableValueFromList('Pindel', delf1) %> & <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'Pindel', '> 1000bp') %>\\ 
&  MoDIL           & <%= printPredTableValueFromList('MoDIL', delPrecision) %>  & <%= printPredTableValueFromList('MoDIL', delRecall) %> & <%= printPredTableValueFromList('MoDIL', delf1) %> & <%= printPredTableCell(delPreds, exDelPreds, 'MoDIL', '40-100bp') %>  & <%= printPredTableCell(delPreds, exDelPreds, 'MoDIL', '101-250bp') %> &  <%= printPredTableCell(delPreds, exDelPreds, 'MoDIL', '251-500bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'MoDIL', '501-1000bp') %> & <%= printPredTableCell(delPreds, exDelPreds, 'MoDIL', '> 1000bp') %>\\ 
   \hline
\multirow{5}{*}{\begin{sideways}Insertions\end{sideways}} & Total Number &          &           & & <%= totalIns['40-100bp'] %> &  <%= totalIns['101-250bp'] %> & <%= totalIns['251-500bp'] %> &  <%= totalIns['501-1000bp'] %> & <%= totalIns['> 1000bp'] %>\\ 
\cline{2-10}
&  Cloudbreak   &<%= printPredTableValueFromList('Cloudbreak', insPrecisionf) %> & <%= printPredTableValueFromList('Cloudbreak', insRecall) %>  & <%= printPredTableValueFromList('Cloudbreak', insf1) %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '40-100bp') %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '101-250bp') %> &  <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '251-500bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '501-1000bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Cloudbreak', '> 1000bp') %> \\ 
&  Breakdancer & <%= printPredTableValueFromList('Breakdancer', insPrecision) %> & <%= printPredTableValueFromList('Breakdancer', insRecall) %>  & <%= printPredTableValueFromList('Breakdancer', insf1) %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '40-100bp') %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '101-250bp') %> &  <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '251-500bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '501-1000bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Breakdancer', '> 1000bp') %> \\ 
&  Pindel          & <%= printPredTableValueFromList('Pindel', insPrecision) %> & <%= printPredTableValueFromList('Pindel', insRecall) %> &  <%= printPredTableValueFromList('Pindel', insf1) %> & <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '40-100bp') %>  & <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '101-250bp') %> &  <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '251-500bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '501-1000bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'Pindel', '> 1000bp') %>\\ 
&  MoDIL          & <%= printPredTableValueFromList('MoDIL', insPrecision) %> & <%= printPredTableValueFromList('MoDIL', insRecall) %> &  <%= printPredTableValueFromList('MoDIL', insf1) %> & <%= printPredTableCell(insPreds, exInsPreds, 'MoDIL', '40-100bp') %>  & <%= printPredTableCell(insPreds, exInsPreds, 'MoDIL', '101-250bp') %> &  <%= printPredTableCell(insPreds, exInsPreds, 'MoDIL', '251-500bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'MoDIL', '501-1000bp') %> & <%= printPredTableCell(insPreds, exInsPreds, 'MoDIL', '> 1000bp') %>\\ 
\hline
\end{tabular}
\end{center}
\caption{The number of simulated deletions and insertions in the 30X diploid chromosome 2 with Venter indels found by each tool at maximum sensitivity, as well as the number of those variants that were discovered exclusively by each tool (in parentheses). The total number of variants in each size class in the true set of deletions and insertions is shown in the first row of each section.}
\label{chr2DeletionAndInsertionPredsMaxSensitivity}
\end{table}