### 2014-04-30 manuscript/combinedRepmaskMultipleMappings.table.brew.tex
tables

\begin{table}
\begin{center}
\begin{tabular}{r|r|rr|rr}
  \cline{2-6}
 & & \multicolumn{2}{c}{Simulated Data} & \multicolumn{2}{c}{NA18507} \\
\hline
 & & Non-repeat & Repeat  & Non-repeat & Repeat \\ 
\multirow{9}{*}{\begin{sideways}Deletions\end{sideways}} & Total Number & <%= chr2DelTotals['0'] %> & <%= chr2DelTotals['1'] %> & <%= NA18507DelTotals['0'] %> & <%= NA18507DelTotals['1'] %> \\ 
  \cline{2-6}
&  Cloudbreak  & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'Cloudbreak', '0') %> & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'Cloudbreak', '1') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'Cloudbreak', '0') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'Cloudbreak', '1') %> \\ 
&  Cloudbreak-GEM  & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'Cloudbreak-GEM', '0') %> & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'Cloudbreak-GEM', '1') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'Cloudbreak-GEM', '0') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'Cloudbreak-GEM', '1') %> \\ 
&  BreakDancer & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'Breakdancer', '0') %> & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'Breakdancer', '1') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'Breakdancer', '0') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'Breakdancer', '1') %> \\
&  GASVPro     & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'GASVPro', '0') %> & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'GASVPro', '1') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'GASVPro', '0') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'GASVPro', '1') %> \\
&  DELLY-RP       & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'DELLY-RP', '0') %> & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'DELLY-RP', '1') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'DELLY-RP', '0') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'DELLY-RP', '1') %> \\
&  DELLY-SR       & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'DELLY-SR', '0') %> & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'DELLY-SR', '1') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'DELLY-SR', '0') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'DELLY-SR', '1') %> \\
&  Pindel      & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'Pindel', '0') %> & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'Pindel', '1') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'Pindel', '0') %> & <%= printPredTableCell(NA18507DelPreds, NA18507ExDelPreds, 'Pindel', '1') %> \\ 
&  MoDIL      & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'MoDIL', '0') %> & <%= printPredTableCell(chr2DelPreds, chr2ExDelPreds, 'MoDIL', '1') %> & NA & NA \\ 
   \hline
\multirow{6}{*}{\begin{sideways}Insertions\end{sideways}} & Total Number & <%= chr2InsTotals['0'] %> & <%= chr2InsTotals['1'] %> & <%= NA18507InsTotals['0'] %> & <%= NA18507InsTotals['1'] %> \\ 
\cline{2-6}
&  Cloudbreak  & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'Cloudbreak', '0') %> & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'Cloudbreak', '1') %> & <%= printPredTableCell(NA18507InsPreds, NA18507ExInsPreds, 'Cloudbreak', '0') %> & <%= printPredTableCell(NA18507InsPreds, NA18507ExInsPreds, 'Cloudbreak', '1') %> \\ 
&  Cloudbreak-GEM  & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'Cloudbreak-GEM', '0') %> & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'Cloudbreak-GEM', '1') %> & <%= printPredTableCell(NA18507InsPreds, NA18507ExInsPreds, 'Cloudbreak-GEM', '0') %> & <%= printPredTableCell(NA18507InsPreds, NA18507ExInsPreds, 'Cloudbreak-GEM', '1') %> \\ 
&  Breakdancer & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'Breakdancer', '0') %> & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'Breakdancer', '1') %> & <%= printPredTableCell(NA18507InsPreds, NA18507ExInsPreds, 'Breakdancer', '0') %> & <%= printPredTableCell(NA18507InsPreds, NA18507ExInsPreds, 'Breakdancer', '1') %> \\
&  Pindel      & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'Pindel', '0') %> & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'Pindel', '1') %> & <%= printPredTableCell(NA18507InsPreds, NA18507ExInsPreds, 'Pindel', '0') %> & <%= printPredTableCell(NA18507InsPreds, NA18507ExInsPreds, 'Pindel', '1') %> \\ 
&  MoDIL      & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'MoDIL', '0') %> & <%= printPredTableCell(chr2InsPreds, chr2ExInsPreds, 'MoDIL', '1') %> & NA & NA \\ 
\hline
\end{tabular}
\end{center}
\caption{Detected deletions and insertions on the simulated and NA18507 data sets identified by each tool, broken down by whether the deletion overlaps with a RepeatMasker-annotated element. For all calls the maximum sensitivity cutoff was used.}
\label{predsByRepmask}
\end{table}